---
title: "Eelgrass Flowering Phenology"
author: "Forest Schenck"
date: "5/25/2023"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

#this is jill's 2nd test!
#Packages Install
```{r}
#only do this if these packages aren't already present on your local drive
install.packages("pillar")
install.packages("tidyr")
install.packages("dplyr")
install.packages("lubridate")
install.packages("ggplot2")
install.packages("mgcv")
```

#Packages Turn-on
```{r}
#Data Carpentry packages
library(pillar)
library(tidyr)
library(dplyr)
library(lubridate)

#Plotting packages
library(ggplot2)
library(glue) #for plot function code; plot title function
library(purrr) #for plot function code; map function

#GAMs Package
library(mgcv)
```

#FS Data upload
##EG 2024
```{r}
d.flw.stage.2024 <- read.csv("~/Downloads/whoi.eelgrass.flw.stage.data.csv")

d.flw.den.2024 <- read.csv("~/Downloads/whoi.eelgrass.density.data.csv")
```

##Data upload Hobos (temp/light) 2024
```{r}
d.t.AQ <- read.csv("~/Downloads/21177827_AQ.csv")

d.t.CB <- read.csv("~/Downloads/21258498_CB.csv")

d.t.CI <- read.csv("~/Downloads/20536645_CI.csv")

d.t.PC <- read.csv("~/Downloads/21177829_PC.csv")

d.t.DC <- read.csv("~/Downloads/21258490_DC.csv")

d.t.LH <- read.csv("~/Downloads/10942817_LH.csv")

d.t.SH <- read.csv("~/Downloads/21177828_SH.csv")

```

#Hobo Temp and Light 2024
##Data carpentry
```{r}
d.t.AQ <- d.t.AQ %>%
  mutate(site.id = "AQ")

d.t.CB <- d.t.CB %>%
  mutate(site.id = "CB")

d.t.CI <- d.t.CI %>%
  mutate(site.id = "CI")

d.t.DC <- d.t.DC %>%
  mutate(site.id = "DC")

d.t.LH <- d.t.LH %>%
  mutate(site.id = "LH")

d.t.PC <- d.t.PC %>%
  mutate(site.id = "PC")

d.t.SH <- d.t.SH %>%
  mutate(site.id = "SH")

str(d.t.SH)

d.t.2024 <- bind_rows(d.t.AQ, d.t.CB, d.t.CI, d.t.DC, d.t.LH, d.t.PC, d.t.SH) %>%
  mutate(date.time = mdy_hm(date.time)) %>%
  filter(date.time > "2024-05-15 19:45") %>%
  filter(date.time < "2024-09-29 20:00") %>%
  mutate(j = as.POSIXlt(date.time)$yday) %>% #calculating julian day
  rename(light = Intensity.lum.per.sqft)

str(d.t.2024)

GDD.2024 <- d.t.2024 %>%
  group_by(site.id, j) %>%
  summarise(max.t = max(temp.C), min.t = min(temp.C), sum.light = sum(light)) %>%
  mutate(GDD = (max.t + min.t)/2 - 8) %>%
  group_by(site.id) %>% 
  mutate(GDD.sum = cumsum(GDD),
         light.sum = cumsum(sum.light))
  
p <-ggplot(data = GDD.2024, aes(x = j, y = GDD.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Day") +
  ylab("GDD") +
  scale_colour_brewer(palette = "PuOr") +
  theme_classic()
  
p

p <-ggplot(data = GDD.2024, aes(x = j, y = light.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Day") +
  ylab("Light") +
  scale_colour_brewer(palette = "PuOr") +
  theme_classic()
  
p

```

#Phenology 2024
##data carpentry (prop)
```{r}
str(d.flw.stage.2024)

d.flw.stage.2024.sum <- d.flw.stage.2024 %>%
  mutate(date.collected = mdy(date.collected)) %>%
  group_by(town, site.id, date.collected, shoot.no) %>%
  summarise(spathe.stage.PS = sum(spathe.stage.PS, na.rm = T), spathe.stage.0 = sum(spathe.stage.0, na.rm = T), spathe.stage.1 = sum(spathe.stage.1, na.rm = T), spathe.stage.2 = sum(spathe.stage.2, na.rm = T), spathe.stage.3 = sum(spathe.stage.3, na.rm = T), spathe.stage.4 = sum(spathe.stage.4, na.rm = T), spathe.stage.5 = sum(spathe.stage.5, na.rm = T), spathe.stage.6 = sum(spathe.stage.6, na.rm = T), spathe.stage.7 = sum(spathe.stage.7, na.rm = T), spathe.stage.unk = sum(spathe.stage.unk, na.rm = T)) %>%
  mutate(no.spathe = spathe.stage.PS + spathe.stage.0 + spathe.stage.1 + spathe.stage.2 + spathe.stage.3 + spathe.stage.4 + spathe.stage.5 + spathe.stage.6 + spathe.stage.7 + spathe.stage.unk,
         stage.PS = spathe.stage.PS/no.spathe,
         stage.0 = spathe.stage.0/no.spathe,
         stage.1 = spathe.stage.1/no.spathe,
         stage.2 = spathe.stage.2/no.spathe,
         stage.3 = spathe.stage.3/no.spathe,
         stage.4 = spathe.stage.4/no.spathe,
         stage.5 = spathe.stage.5/no.spathe,
         stage.6 = spathe.stage.6/no.spathe,
         stage.7 = spathe.stage.7/no.spathe,
         stage.unk = spathe.stage.unk/no.spathe) %>%
  select(town:shoot.no, stage.PS:stage.unk) %>%
  pivot_longer(!town:shoot.no, names_to = "flw.stage", values_to = "prop") %>%
  mutate(flw.stage = as.factor(flw.stage),
         shoot.no = as.factor(shoot.no)) %>%
  mutate(flw.stage = factor(flw.stage, levels=c ('stage.PS', 'stage.0', 'stage.1', 'stage.2', 'stage.3', 'stage.4', 'stage.5', 'stage.6', 'stage.7', 'stage.unk')))

d.flw.stage.2024.sum = as.data.frame(d.flw.stage.2024.sum)


str(d.flw.stage.2024.sum)
```

##Stage Proportion Plot Function (prop)
```{r}
stage.plot <- function(site.code) {
  d.flw.stage.2024.sum %>%
    filter(.data$site.id == .env$site.code) %>%
    ggplot() +
    aes(x = date.collected, y = prop, color = flw.stage) +
    geom_point() +
    geom_smooth(se = FALSE) +
    xlab("Date") +
    ylab("Proportion") +
    scale_colour_brewer(palette = "PuOr", direction = -1) +
    ggtitle(glue("Site: {site.code}")) +
    theme(plot.title.position = "plot") +
    theme_classic()
}

site.code <- c("SH", "CI", "PC", "LH", "PI", "AQ", "WB", "NB", "CB", "DC")

site.plots <- map(site.code, stage.plot)

site.plots[1:10]

```

##data carpentry (mean stage)
```{r}
str(d.flw.stage.2024)

d.flw.stage.2024.mean <- d.flw.stage.2024 %>%
  mutate(date.collected = mdy(date.collected)) %>%
  group_by(town, site.id, date.collected, shoot.no) %>%
  summarise(spathe.stage.PS = sum(spathe.stage.PS, na.rm = T), spathe.stage.0 = sum(spathe.stage.0, na.rm = T), spathe.stage.1 = sum(spathe.stage.1, na.rm = T), spathe.stage.2 = sum(spathe.stage.2, na.rm = T), spathe.stage.3 = sum(spathe.stage.3, na.rm = T), spathe.stage.4 = sum(spathe.stage.4, na.rm = T), spathe.stage.5 = sum(spathe.stage.5, na.rm = T), spathe.stage.6 = sum(spathe.stage.6, na.rm = T), spathe.stage.7 = sum(spathe.stage.7, na.rm = T), spathe.stage.unk = sum(spathe.stage.unk, na.rm = T)) %>%
  mutate(spathe.stage.0 = spathe.stage.PS + spathe.stage.0,
         spathe.stage.6 = spathe.stage.6 + spathe.stage.7,
         no.spathe = + spathe.stage.0 + spathe.stage.1 + spathe.stage.2 + spathe.stage.3 + spathe.stage.4 + spathe.stage.5 + spathe.stage.6,
         stage.0 = (spathe.stage.0/no.spathe)*0,
         stage.1 = (spathe.stage.1/no.spathe)*1,
         stage.2 = (spathe.stage.2/no.spathe)*2,
         stage.3 = (spathe.stage.3/no.spathe)*3,
         stage.4 = (spathe.stage.4/no.spathe)*4,
         stage.5 = (spathe.stage.5/no.spathe)*5,
         stage.6 = (spathe.stage.6/no.spathe)*6,
         stage.sum = stage.0 + stage.1 + stage.2 + stage.3 + stage.4 +stage.5 + stage.6) %>%
  select(town:shoot.no, stage.sum) %>%
  mutate(j = as.POSIXlt(date.collected)$yday)

d.flw.stage.2024.mean = as.data.frame(d.flw.stage.2024.mean)


str(d.flw.stage.2024.mean)
```

##Stage Proportion Plot Function (mean stage)
```{r}
p <-ggplot(data = d.flw.stage.2024.mean, aes(x = j, y = stage.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Day") +
  ylab("Mean Stage Value") +
  scale_colour_brewer(palette = "PuOr") +
  geom_hline(yintercept = 4) +
  theme_classic()
  
p

#zoom plot
p <-ggplot(data = d.flw.stage.2024.mean, aes(x =j, y = stage.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Date") +
  ylab("Mean Stage Value") +
  scale_colour_brewer(palette = "PuOr") +
  geom_hline(yintercept = 4) +
  coord_cartesian(ylim = c(3.5,4.5), xlim = c(190,225)) +
  scale_x_continuous(breaks = seq(190,230, by = 2)) +
  theme_classic()
p
```

##Plot mean stage prop by GDD
```{r}
d.flw.stage.2024.mean.GDD <- left_join(d.flw.stage.2024.mean, GDD.2024, by = c("site.id", "j"))

d.flw.stage.2024.mean.GDD <- d.flw.stage.2024.mean.GDD %>%
  filter(!is.na(GDD.sum))

p <-ggplot(data = d.flw.stage.2024.mean.GDD, aes(x = GDD.sum, y = stage.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("GDD") +
  ylab("Mean Stage Value") +
  scale_colour_brewer(palette = "PuOr") +
  geom_hline(yintercept = 4) +
  theme_classic()
  
p


p <-ggplot(data = d.flw.stage.2024.mean.GDD, aes(x = j, y = stage.sum, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Day") +
  ylab("Mean Stage Value") +
  scale_colour_brewer(palette = "PuOr") +
  geom_hline(yintercept = 4) +
  theme_classic()
  
p
```

#Seed Production Per Shoot 2024
##data carpentry
```{r}
str(d.flw.stage.2024)

d.seed.production.2024.sum <- d.flw.stage.2024 %>%
  mutate(date.collected = mdy(date.collected)) %>%
  rowwise() %>%
  mutate(no.seeds.total = sum(no.seeds, no.ovules.w.stigma, na.rm = T)) %>%
  group_by(town, site.id, date.collected, shoot.no) %>%
  summarise(no.seeds.total = sum(no.seeds.total, na.rm = T))

d.seed.production.2024.sum = as.data.frame(d.seed.production.2024.sum)


str(d.seed.production.2024.sum)
```

##Seed production per Shoot Plot Function
```{r}

seeds.per.shoot.plot <- function(site.code) {
  d.seed.production.2024.sum %>%
    filter(.data$site.id == .env$site.code) %>%
    ggplot() +
    aes(x = date.collected, y = no.seeds.total) +
    geom_point() +
    geom_smooth(se = FALSE) +
    xlab("Date") +
    ylab("Seeds per shoot") +
    ggtitle(glue("Site: {site.code}")) +
    theme(plot.title.position = "plot") +
    theme_classic()
}

site.code <- c("SH", "CI", "PC", "LH", "AQ", "WB", "CB", "DC")

site.plots <- map(site.code, seeds.per.shoot.plot)

site.plots[1:8]



p <-ggplot(data = d.seed.production.2024.sum, aes(x = date.collected, y = no.seeds.total, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Date") +
  ylab("Seeds per shoot") +
  scale_colour_brewer(palette = "PuOr") +
  theme_classic()
  
p

```


#Flw Density 2024
##data carpentry
```{r}
str(d.flw.den.2024)

d.flw.den.2024.clean <- d.flw.den.2024 %>%
  mutate(date.collected = mdy(date.collected)) %>%
  select(date.collected, site.id, quadrat.no, shoot.type, shoot.no) %>%
  spread(key = shoot.type, value = shoot.no) %>%
  filter(reproductive > 0 | vegetative > 0) %>% #only plots where eelgrass is present
  mutate(flw.m.sq = reproductive * 4)

```

##Flw Density Plot Function
```{r}
p <-ggplot(data = d.flw.den.2024.clean, aes(x = date.collected, y = flw.m.sq, color = site.id)) +
  geom_point() +
  geom_smooth() +
  xlab("Date") +
  ylab("Repros per m2") +
  scale_colour_brewer(palette = "PuOr") +
  theme_classic()
  
p
```


#Seed Density 2024
##Data Carpentry
```{r}
#Summarise to mean reproductive shoot density in plots where eelgrass present for each monitoring date
d.flw.den.2024.sd <- d.flw.den.2024 %>%
  mutate(date.collected = mdy(date.collected)) %>%
  select(date.collected, site.id, latitude, longitude, quadrat.no, shoot.type, shoot.no) %>%
  spread(key = shoot.type, value = shoot.no) %>%
  filter(reproductive > 0 | vegetative > 0) %>% #only plots where eelgrass is present
  mutate(flw.m.sq = reproductive * 4) %>%
  group_by(date.collected, site.id, latitude, longitude) %>%
  summarise(mean.flw.m.sq = mean(flw.m.sq))

#Summarise to mean seeds per flw for each monitoring date
d.seed.production.2024.sd <- d.seed.production.2024.sum %>%
  filter(!is.na(shoot.no)) %>%
  group_by(site.id, date.collected) %>%
  summarise(mean.seeds = mean(no.seeds.total))

#merge flw density and seed data sets, include dates where density measured but no seeds
d.seed.den.2024 <- merge(d.flw.den.2024.sd, d.seed.production.2024.sd, by = c("site.id","date.collected"), all.x = T)

#Cleaning up merged seed production data set
d.seed.den.2024.f <- d.seed.den.2024 %>%
  mutate(mean.seeds = replace(mean.seeds, is.na(mean.seeds), 0)) %>% #replacing dates with no seeds to 0
  mutate(seeds.m.sq = mean.flw.m.sq*mean.seeds, #calculating seeds per m2
         j = as.POSIXlt(date.collected)$yday) #calculating julian day

d.seed.den.2024.GDD <- left_join(d.seed.den.2024.f, GDD.2024, by = c("site.id", "j")) #Adding GDD to seed production data set

d.seed.den.2024.GDD <- d.seed.den.2024.GDD %>%
  filter(!is.na(GDD.sum))
  

```

##Plot by Date
```{r}
p <-ggplot(data = d.seed.den.2024.f, aes(x = j, y = seeds.m.sq+1, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Date") +
  ylab("Seeds per m2") +
  scale_colour_brewer(palette = "PuOr") +
  scale_y_continuous(trans = 'log10') +
  theme_classic()
  
p


p <-ggplot(data = d.seed.den.2024.f, aes(x =j, y = seeds.m.sq+1, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("Julian Date") +
  ylab("Mean Stage Value") +
  scale_colour_brewer(palette = "PuOr") +
  geom_hline(yintercept = 4) +
  coord_cartesian(ylim = c(90,7000), xlim = c(170,240)) +
  scale_y_continuous(trans = 'log10') +
  scale_x_continuous(breaks = seq(170,240, by = 2)) +
  theme_classic()
p

```

##Plot by GDD
```{r}
p <-ggplot(data = d.seed.den.2024.GDD, aes(x = GDD.sum, y = seeds.m.sq+1, color = site.id)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlab("GDD") +
  ylab("Seeds per m2") +
  scale_colour_brewer(palette = "PuOr") +
  scale_y_continuous(trans = 'log10') +
  theme_classic()
  
p
```


#GAM Trials Seed Density
##GAM J-day
```{r}
# SELECT AN ERROR DISTRIBUTION
  plot(density(d.seed.den.2024.f$seeds.m.sq))

  m1<-gam(seeds.m.sq~1, data=d.seed.den.2024.f)   
  m2<-gam(seeds.m.sq~1, data=d.seed.den.2024.f, family='nb') #NEGATIVE BINOMIAL
  m3<-gam(seeds.m.sq~1, data=d.seed.den.2024.f, family='tw') #TWEEDIE
  AIC(m1,m2,m3)
  fam<-'nb'

  
#GAM seeds by date
mod.j <- gam(seeds.m.sq~s(j, bs = 'cc'), data=d.seed.den.2024.f, family=fam, select=TRUE, method="REML") 
summary(mod.j)
plot(mod.j,
     xlim = c(127,295))

#GAM seeds by latitude
mod.lat <- gam(seeds.m.sq ~ s(latitude), data = d.seed.den.2024.f, family = fam, select = TRUE, method = "REML")
summary(mod.lat)
plot(mod.lat)

#GAM seeds by latitude and date
mod.j.lat <- gam(seeds.m.sq~te(j, latitude, bs = c('cc', 'ts'))
                 , data = d.seed.den.2024.f, family = fam, select = TRUE, method = "REML")

summary(mod.j.lat)
plot(mod.j.lat)

```

##GAM with GDD and total light
```{r}
# SELECT AN ERROR DISTRIBUTION
  plot(density(d.seed.den.2024.GDD$seeds.m.sq))

  m1<-gam(seeds.m.sq~1, data=d.seed.den.2024.GDD)   
  m2<-gam(seeds.m.sq~1, data=d.seed.den.2024.GDD, family='nb') #NEGATIVE BINOMIAL
  m3<-gam(seeds.m.sq~1, data=d.seed.den.2024.GDD, family='tw') #TWEEDIE
  AIC(m1,m2,m3)
  fam<-'nb'

  
#GAM seeds by date
mod.j <- gam(seeds.m.sq~s(j, bs = 'cc'), data=d.seed.den.2024.GDD, family=fam, select=TRUE, method="REML") 
summary(mod.j)
plot(mod.j,
     xlim = c(148,256))

#GAM seeds by latitude
mod.lat <- gam(seeds.m.sq ~ s(latitude, k = 5), data = d.seed.den.2024.GDD, family = fam, select = TRUE, method = "REML")
summary(mod.lat)
plot(mod.lat)

#GAM seeds by GDD
mod.GDD <- gam(seeds.m.sq ~ s(GDD.sum), data = d.seed.den.2024.GDD, family = fam, select = TRUE, method = "REML")
summary(mod.GDD)
plot(mod.GDD)

#GAM seeds by light
mod.light <- gam(seeds.m.sq ~ s(light.sum), data = d.seed.den.2024.GDD, family = fam, select = TRUE, method = "REML")
summary(mod.light)
plot(mod.light)

#GAM seeds by GDD, light independent
mod.GDD.light.ind<-gam(seeds.m.sq~s(GDD.sum) #GDD
               +s(light.sum) #Light
               ,data=d.seed.den.2024.GDD,family=fam,select=TRUE,method="REML")
summary(mod.GDD.light.ind)
plot(mod.GDD.light.ind)

#GAM seeds by GDD x light interaction
mod.GDD.light <- gam(seeds.m.sq ~ te(GDD.sum, light.sum), data = d.seed.den.2024.GDD, family = fam, select = TRUE, method = "REML")
summary(mod.GDD.light)
plot(mod.GDD.light)


#GAM Full GDD, light model
mod.full<-gam(seeds.m.sq~s(GDD.sum) #GDD
              +s(light.sum) #Light
              +s(j, bs = 'cc') #julian day
              +s(latitude, k = 5)
              +te(GDD.sum, light.sum) #light X GDD interaction
               ,data=d.seed.den.2024.GDD,family=fam,select=TRUE,method="REML")
summary(mod.full)
plot(mod.full)
qq.gam(mod.full) #IS THE ERROR DISTRIBUTION APPROPRIATE?
gam.check(mod.full) #CHECK TO SEE IF SMOOTHERS HAVE BEEN GIVEN SUFFICIENT EDF

```

#GAM Trials Stage Proportion
##GAM with GDD
```{r}
d.flw.stage.2024.GAM <- d.flw.stage.2024.mean.GDD %>%
  group_by(site.id, j) %>%
  summarise(stage.sum = mean(stage.sum), GDD.sum = mean(GDD.sum), light.sum = mean(light.sum))

# SELECT AN ERROR DISTRIBUTION
  plot(density(d.flw.stage.2024.GAM$stage.sum))

  m1<-gam(stage.sum~1, data=d.flw.stage.2024.GAM)   
  m2<-gam(stage.sum~1, data=d.flw.stage.2024.GAM, family='nb') #NEGATIVE BINOMIAL
  m3<-gam(stage.sum~1, data=d.flw.stage.2024.GAM, family='tw') #TWEEDIE
  AIC(m1,m2,m3)

#GAM stage by GDD
mod.GDD <- gam(stage.sum ~ s(GDD.sum), data = d.flw.stage.2024.GAM, select = TRUE, method = "REML")
summary(mod.GDD)
plot(mod.GDD)

qq.gam(mod.GDD) #IS THE ERROR DISTRIBUTION APPROPRIATE?
gam.check(mod.GDD) #CHECK TO SEE IF SMOOTHERS HAVE BEEN GIVEN SUFFICIENT EDF

#GAM stage by j-day
mod.j <- gam(stage.sum ~ s(j, bs = 'cc'), data = d.flw.stage.2024.GAM, select = TRUE, method = "REML")
summary(mod.j)
plot(mod.j,
     xlim = c(148,256))

#GAM stage by light
mod.light <- gam(stage.sum ~ s(light.sum), data = d.flw.stage.2024.GAM, select = TRUE, method = "REML")
summary(mod.light)
plot(mod.light)


#GAM stage by GDD x light interaction
mod.GDD.light <- gam(stage.sum ~ te(GDD.sum, light.sum), data = d.flw.stage.2024.GAM, select = TRUE, method = "REML")
summary(mod.GDD.light)
plot(mod.GDD.light)

#GAM Full GDD, light model
mod.full<-gam(stage.sum~s(GDD.sum) #GDD
              +s(light.sum) #Light
              +s(j, bs = 'cc') #julian day
              +te(GDD.sum, light.sum) #light X GDD interaction
               ,data=d.flw.stage.2024.GAM,select=TRUE,method="REML")
summary(mod.full)
plot(mod.full)
qq.gam(mod.full) #IS THE ERROR DISTRIBUTION APPROPRIATE?
gam.check(mod.full) #CHECK TO SEE IF SMOOTHERS HAVE BEEN GIVEN SUFFICIENT EDF
```

